### 2023-02-27 code/analysis_code/1_EDA_sex.Rmd
comments
changes

```{r message=FALSE}
library(here)
library(GGally)
library(skimr)
library(ggfortify)
library(car)
library(bestNormalize)
library(MASS)
library(tidyverse)
library(readxl)
```
Loading data
```{r}
sex_df <- readRDS(here("data/processed_data/sex_df_cleaned.rds"))
```

@@

par(mfrow=c(1,3))
plot(lm(sex_df_DR$BOX ~ 1))
```
Unsuccessful.

@@

par(mfrow=c(1,3))
plot(lm(sex_df_SMR$BOX ~ 1))
```
The Box Cox transformation was not able to normalize the `Standardized Mortality Ratio`. We will move on the Yeo Johnson transformation

We will perform Yeo Johnson transformation on `Death Rate (estimated)` and `Standardized Mortality Ratio`.We will use `yeojohnson` from `BestNormalize` package to extract lambda and `yeo_johnson` from `car` to compute the transformation for each observation.
```{r}
yj_DR <- yeojohnson(sex_df_SMR$`Death Rate (estimated)`)
yj_DR <- yj_DR$lambda

@@

par(mfrow=c(1,3))
plot(lm(sex_df$YJ ~ 1))
```
Unable to normalized the distribution `Death Rate (estimated)` with the Yeo Johnson transformation.

@@

Plotting correlation matrix
```{r message=FALSE, warning=FALSE}
X <- sex_df[,c(9:10, 16:25)]
ggpairs(X)
```

```{r}
saveRDS(sex_df_ROE, here("data/process_data/sex_new.rds"))
```

>>>>

```{r message=FALSE}
library(here)
library(GGally)
library(skimr)
library(ggfortify)
library(MASS)
library(tidyverse)
library(readxl)
```
Loading data
```{r}
sex_df <- readRDS(here("data/processed_data/sex_df_cleaned.rds"))
```

@@

par(mfrow=c(1,3))
plot(lm(sex_df_DR$BOX ~ 1))
```
Unsuccessful at normalizing distribution of Death Rate, and overall worse results than log transformation, except for the males.

@@

par(mfrow=c(1,3))
plot(lm(sex_df_SMR$BOX ~ 1))
```
The Box Cox transformation was not technically able to normalize the `Standardized Mortality Ratio`. We will move on the Yeo Johnson transformation

We will perform Yeo Johnson transformation on `Death Rate (estimated)` and `Standardized Mortality Ratio`.We will use `yeojohnson` from `BestNormalize` package to extract lambda and `yeo_johnson` to compute the transformation for each observation. Code for `yeo_johnson` obtained from [GitHub](https://stats.stackexchange.com/questions/1444/how-should-i-transform-non-negative-data-including-zeros)
```{r}
library(bestNormalize)

yeo_johnson <- function(x, lambda) {

    eps <- .000001
    not_neg <- which(x >= 0)
    is_neg  <- which(x < 0)

    not_neg_trans <- function(x, lambda) {
        if (abs(lambda) < eps) log(x + 1)
        else ((x + 1) ^ lambda - 1) / lambda
    }

    neg_trans <- function(x, lambda) {
        if (abs(lambda - 2) < eps) - log(-x + 1)
        else - ((-x + 1) ^ (2 - lambda) - 1) / (2 - lambda)
    }

    x[not_neg] <- not_neg_trans(x[not_neg], lambda)

    x[is_neg] <- neg_trans(x[is_neg], lambda)

    return(x)

}

```{r}
yj_DR <- yeojohnson(sex_df_DR$`Death Rate (estimated)`)
yj_DR <- yj_DR$lambda

@@

par(mfrow=c(1,3))
plot(lm(sex_df$YJ ~ 1))
```
Unable to normalized the distribution `Death Rate (estimated)` with the Yeo Johnson transformation, but improved for overall and males compared to log transformation (not for females).

@@

Plotting correlation matrix
```{r message=FALSE, warning=FALSE}
X <- sex_df[,c(9:10, 16:25)]
ggpairs(X)
```
Let's save the dataframe for ROE since this one was able to be normalized after Box Cox transformation.
```{r}
saveRDS(sex_df_ROE, here("data/processed_data/sex_new.rds"))
```
